---
title: "Analysis of World Population"
description: |
  This website would give you an idea about the current world population and different indices based upon population.
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(distill)
library(rvest)
library(tidyverse)
library(ggthemes)
library(plotly)
library(scales)
library(janitor)

# library()
# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

# Learn more about publishing to GitHub Pages at:
# https://rstudio.github.io/distill/publish_website.html#github-pages

pop_den_grow <- read.csv('Population, Growth and Density Data.csv')
loan <- read.csv('Loan Progression.csv')
health <- read.csv('Life Expectancy Data.csv') |>
    filter(Year == 2014)
economics_life <- merge(x = pop_den_grow, y = health, by = "Country")
economics_life <- merge(x = economics_life, y = loan, by = "Country")
economics_life <- economics_life |>
  mutate(GrowthRate = (GrowthRate - 1))
# View(economics_life)
```

## Population and Growth Rate

The relation between population of a country is depicted in the graph below. Key finding from this graph is "**Growth rate is inversely proportional to income category."**

```{r}

# Population and Density
economics_life |>
  select(Country, pop2022, Density, Continent) |>
  ggplot(aes(Density, pop2022, color = Continent)) +
  geom_point()+
    theme_classic()+
    scale_x_log10(labels = comma)+
    scale_y_log10(labels = comma)+
    #geom_smooth(se=F)+
    labs(title = "Relation between population and density",
         subtitle = "Countries with smaller population tend to have higher population density",
         x = "Density",
         y = "Population",
         color = "Continent")

```

The key takeaways from this graph:

-   Most countries tend to have density around 100 to 500 per square kilometer.

-   Countries with **small** population tend to have **very** **high** population density.

-   Countries with **large** population tend to have **low** population density.

## Land Area and Density

The relation between growth rate of a country and land area is depicted in the graph below. Key finding from this graph is "**Density is inversely proportional to land area."**

```{r}
economics_life |>
  ggplot(aes(Density, Area, color = Continent)) +
  geom_point()+
    theme_classic()+
    scale_x_log10(labels = comma)+
    scale_y_log10(labels = comma)+
    labs(title = "Relation between land area and density",
         subtitle = "Countries with smaller land area tend to have higher population density",
         x = "Density",
         y = "Population",
         color = "Continent")

```

The key takeaways from this graph:

-   Most countries tend to have density around 100 to 500 per square kilometer.

-   Countries with **small** land area tend to have **very** **high** population density.

-   Countries with **large** land area tend to have **low** population density.

## Income Group and Growth Rate

The relation between growth rate of a country and per capita income is depicted in the graph below. Key finding from this graph is "Growth rate is inversely proportional to per capita income."

```{r}
# View(economics_life)
economics_life |>
    select(GrowthRate, income.group, Region, Continent) |>
    ggplot(aes(y = GrowthRate, x = income.group, color = Continent))  +
    geom_point() +
    theme_classic()+
    scale_y_continuous(labels = percent) +
    labs(title = "Growth rate across the globe",
         subtitle = "Countries with low income tend to have high growth rate",
         x = "Income group",
         y = "Growth rate",
        color = "Continent")
```

```{r}

economics_life |>
    select(GrowthRate, income.group, Region, Continent) |>
    ggplot(aes(y = GrowthRate, x = Continent, color = income.group)) +
    geom_point()+
    theme_classic()+
    scale_y_continuous(labels = percent) +
    labs(title = "Growth rate across the globe",
         subtitle = "Countries with low income tend to have high growth rate",
         x = "Continent",
         y = "Growth rate",
         color = "Income group")
```

```{r}

economics_life |>
    select(GrowthRate, Status, Region) |>
    ggplot(aes(y = GrowthRate, x = Status, color = Region)) +
    geom_point()+
    theme_classic()+
    scale_y_continuous(labels = percent) +
    labs(title = "Growth rate across the globe",
         subtitle = "Countries with low income tend to have high growth rate",
         x = "Economic Progress",
         y = "Growth rate",
         color = "Region")
```

The key takeaways from this graph:

-   Most countries tend to have density around 100 to 500 per square kilometer.

-   Countries with **low income** fall into the **developing category** and have **very high growth rate**.

-   Countries with **high income** fall into the **developed category** and have **very low growth rate**. Generally those countries have a growth rate lying between **-1%** and **1%**.

## Developed and Developing Nations

```{r}


economics_life |>
    ggplot(aes(y = Status)) +
    theme_classic()+
    geom_bar(aes(fill = Region)) +
    labs(title = "Developed and developing countries around the world",
         subtitles = "The amount of developing countries surpass developed countries by landslide",
         x = "Economic Progress",
         y = "Count",
         color = "Continent")

economics_life |>
    ggplot(aes(y = Continent)) +
    theme_classic()+
    geom_bar(aes(fill = Status))+
    labs(title = "Developed and developing countries around the world",
         subtitles = "Except Europe all the continents are dominated by developing countries",
         x = "Continent",
         y = "Count",
         color = "Economic Progress")


economics_life |>
    ggplot(aes(y = Region)) +
    theme_classic()+
    geom_bar(aes(fill = Status))+
    labs(title = "Developed and developing countries around the world",
         subtitles = "Except Europe all the continents are dominated by developing countries",
         x = "Continent",
         y = "Count",
         color = "Economic Progress")



economics_life |>
    ggplot(aes(y = income.group)) + 
    theme_classic()+
    geom_bar(aes(fill = Continent)) +
    
    labs(title = "Developed and developing countries around the world",
         subtitles = "Most high Income countries in Europe, low income are in Africe, and middle earners are diverse",
         x = "Income Group",
         y = "Count",
         color = "Continent")
```

Key takeaways from the graph are listed below:

-   **Europe** is dominated by developed countries, but the rest are filled with developing countries.

-   Developing countries are more common that developed countries.

-   **South America**, **North America** and **Africa** is filled with mostly developing countries.

-   **Asia** and **Oceania** is consists of both developing and developed countries.

## Lending Category and Economy

**IDA** stands for **International Development Association**, which is one of the largest and most effective platforms for fighting extreme poverty in the world's poorest countries. On the other hand, **IBRD** stands for the **International Bank for Reconstruction and Development,** commonly referred as **World Bank**. **IBRD** assists the development of its member nation's territories, promotes and supplements private foreign investment and promoting long-range balance.

```{r}

economics_life |>
    ggplot(aes(Lending.category)) +
    theme_classic()+
    geom_bar(aes(fill = Continent)) +
    labs(title = "Lending Category Analysis",
         subtitles = "Most High Income countries are located in Europe and most low income are in Africe, and Middle earners are extremelty diverse",
         x = "Income Group",
         y = "Count",
         color = "Continent")
```

```{r}

economics_life |>
    ggplot(aes(Continent)) +
    theme_classic()+
    geom_bar(aes(fill = Lending.category)) +
    labs(title = "Developed and developing countries around the world",
         subtitles = "Most High Income countries are located in Europe and most low income are in Africe, and Middle earners are extremelty diverse",
         x = "Income Group",
         y = "Count",
         color = "Continent")


economics_life |>
    ggplot(aes(Status)) +
    theme_classic()+
    geom_bar(aes(fill = Lending.category)) +
    labs(title = "Developed and developing countries around the world",
         subtitles = "Most High Income countries are located in Europe and most low income are in Africe, and Middle earners are extremelty diverse",
         x = "Income Group",
         y = "Count",
         color = "Continent")


```

Key takeaways from the graph are listed below:

-   Most developing **countries** are located in **Africa**. So these nations have to depend a lot on loans from **IDA** to develop the quality of life.

-   Europe is the most developed continent, therefore, most European nations don't need any loans to operate their economy.

-   Asia consists of both developed and developing nations. So there's a diversity among Asian nations.
